Automated Classification of TA Comments on Student Writing

### 2019-02-21 background.Rmd
update

---
title: "Background for the Project"
bibliography: library.bib
biblio-style: apalike
---

@@

## Context for This Project
From prior studies (REFs), we know students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead. To that end, we train graduate teaching assistants to use a **global issues first, specific issues later** approach to correcting undergraduate technical writing. We ask TAs to spend a limited block of time writing fewer comments that focus on broader issues. Ideally, TAs provide no more than 3-5 comments per text page, and comments should encourage student reflection. As students' writing improves, the TA can focus on smaller issues. Our bins-based report grading protocol reinforces this strategy.

@@

## Prior Work
In 2018, we began trying to answer this question by analyzing TAs' comments from student papers directly. We extracted and tabulated all comments from approximately 500 pairs of reports. After initial data cleaning (removing duplicates, splitting comments addressing multiple independent writing issues, and removing comments lacking any context for judging intent), we had a working dataset of ~11,000 separate comments.

@@

## The Central Problem of This Project
Coding TA comments this way is informative but doing so by hand is unsustainable. Developing the codebook then rating the initial set of 11,000 comments required nearly 2 months of investigator effort. Even with the codebook in place, individual coders need training to achieve sufficient accuracy, and there is significant risk of "coding drift" over time.

@@

>>>>

---
title: "Background"
bibliography: library.bib
biblio-style: apalike
---

@@

####\  
##Context for This Project
From prior studies (REFs), we know students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead. To that end, we train graduate teaching assistants to use a **global issues first, specific issues later** approach to correcting undergraduate technical writing. We ask TAs to spend a limited block of time writing fewer comments that focus on broader issues. Ideally, TAs provide no more than 3-5 comments per text page, and comments should encourage student reflection. As students' writing improves, the TA can focus on smaller issues. Our bins-based report grading protocol reinforces this strategy.

@@

###Prior Work
In 2018, we began trying to answer this question by analyzing TAs' comments from student papers directly. We extracted and tabulated all comments from approximately 500 pairs of reports. After initial data cleaning (removing duplicates, splitting comments addressing multiple independent writing issues, and removing comments lacking any context for judging intent), we had a working dataset of ~11,000 separate comments.

@@

##Central Problem
Coding TA comments this way is informative but doing so by hand is unsustainable. Developing the codebook then rating the initial set of 11,000 comments required nearly 2 months of investigator effort. Even with the codebook in place, individual coders need training to achieve sufficient accuracy, and there is significant risk of "coding drift" over time.

@@

##Work Plan

1.  Compile initial dataset, and de-identify both TAs and students by replacing names with anonymous IDs.
2.  Export TA comments, metadata from student reports as a "tidy" dataframe.
3.  Explore comment data structure independently of classification to identify useful elements for feature engineering.
4.  Write the initial Naive Bayes workflow for supervised text classification.
5.  Assign TA comments to categories defined in our existing codebook, then compare outcomes from hand-coding versus automated classification.
6.  Using features (#3) and comparison data from initial classification protocol (#4) as a guide, test iterations of the classifier method and optimize. 
7.  Apply the optimized classifier to the original comment dataset, identify all comments that are being classified incorrectly, and look for patterns.
8.  Write a small set of rule or REGEX pattern-based searches that can identify comments which are more likely to be classified incorrectly. 
9.  Append rules/pattern-based pre-screening to the optimized Naive Bayes classifier.
10.  Validate this vertical Ensemble Classifier on a subset of ~2000 TA comments extracted from a non-overlapping set of student lab reports.

### 2019-02-21 index.Rmd
update

## Overview
I am using classroom research and data science to better understand how students develop technical writing skills in large-enrollment biology courses.

@@

#### \   
## Prior Work
In Summer 2018, we extracted and tabulated ~11,000 comments made by teaching assistants (TA) from >500 pairs (initial submissions and revised version) of short lab reports written by undergraduates in 3 different general biology lab courses during a single semester. Using an iteratively generated codebook, each of the comments was separately classified in two categories:

@@

Classifying TA comments this way is informative, but doing so by hand is impractical: 10-12,000 comments need to be scored EACH semester. There also is a significant risk of "coding drift" over time, reducing accuracy.

@@

##Specific Project Goals
My project for the **Faculty Learning Community for R** was to create an automated comment classifier that could generate two-axis contingency tables comparable to the one generated using hand-coded data. This automated classifier will used within the larger NSF-funded research project to:

#### \   
## Site Map
Insert a text-based overview and site map of the major pages and branch points here.

>>>>

##Overview
I am using classroom research and data science to better understand how students develop technical writing skills in large-enrollment biology courses.

@@

In Summer 2018, we extracted and tabulated ~11,000 comments made by teaching assistants (TA) from >500 pairs (initial submissions and revised version) of short lab reports written by undergraduates in 3 different general biology lab courses during a single semester. Using an iteratively generated codebook, each of the comments was separately classified in two categories:

@@

#### \   
##Specific Goals
Classifying TA comments this way is informative, but doing so by hand is impractical: 10-12,000 comments need to be scored EACH semester. There also is a significant risk of "coding drift" over time, reducing accuracy.

My project for the **Faculty Learning Community for R** was to create an automated comment classifier that could generate two-axis contingency tables comparable to the one generated using hand-coded data. This automated classifier will used within the larger NSF-funded research project to:

#### \   
##Project Site Map
Insert a text-based overview and site map of the major pages and branch points here.

When appropriate, anchor pages branch to extended supporting documentation.

@@
